---
title: "Lab 3 Intro"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
```

## Setup
1. Go to iclicker.com and record your attendance.
2. Login to the lab computer (please don't use your personal computer).  
3. Navigate to github.com and login.  
4. Find the class repository "datascibiol" (https://github.com/jmledford3115/datascibiol).  
5. Download a copy the class repository to the desktop.  
6. Clone your repository to the desktop.  
7. Copy the files for today's lab from the class repository and paste them into **your** repository.  
8. Open today's lab in RStudio.  

## Review from last time
### *With a partner, discuss the following questions*
1. What are examples of different classes of objects in R?
Some examples of different classes of objects in R include numeric, character, and logical.

2. What is a vector? What is the command to make a vector in R?
A vector is a one-dimensional array of data. The command to make a vector in R is c().

3. What is a data matrix and how does a data matrix relate to vectors?  
A data matrix is a two-dimensional array of data. A data matrix is made up of vectors.

## Warm-up
1. Build a vector that includes the following height measurements for five plants:
plant 1 30.7, plant 2 37.6, plant 3 28.4, plant 4 NA, plant 5 33.2
```{r}
plant_height <- c(30.7, 37.6, 28.4, NA, 33.2)

plant_height
```


2. Build another vector that includes the following mass measurements:
plant 1 4, plant 2 5.2, plant 3 3.7, plant 4 NA, plant 5 4.6
```{r}
plant_mass <- c(4, 5.2, 3.7, NA, 4.6)

plant_mass
```

3. Assemble these vectors into a labeled data matrix with two columns
```{r}
plant_data <- cbind(plant_height, plant_mass)

plant_data
```

4. Calculate the mean for height and mass and add them to the data matrix
```{r}
plant_data <- rbind(plant_data, c(mean(plant_height, na.rm = TRUE), mean(plant_mass, na.rm = TRUE)))

plant_data


```

